
---
title: "Reproduce_Project_2"
author: "Sudabe"
date: "April 16, 2016"
output: html_document
---
First I read the data in r and name the file activity.


```{r}
activity<-read.csv("./activity.csv", header = TRUE)

```


```{r}
library(dplyr)
library(lubridate)
```
I named the file activity.  I also need to use "dplyr" and "lubridate" libraries. 

## What is mean total number of steps taken per day?

### Total number of steps:


```{r}
activity$date=as.character(activity$date)
activity$date=as.Date(activity$date)
activity$day=yday(activity$date)
activity_day=group_by(activity, day)
Total_step_day <- summarise(activity_day, steps_day=sum(steps, na.rm = TRUE ))

```
### Histogram:

```{r}
library(ggplot2)
g=ggplot(Total_step_day, aes(day, steps_day))+geom_histogram(stat = "identity", fill="blue")+ggtitle("Total Number of Steps per Day")+ xlab("Day")+ylab("Total Number of Steps")
plot(g)
```

### Calculate and report the mean and median of the total number of steps taken per day:

```{r}
step_day_stat <- summarise(activity_day, mean_steps_day=mean(steps, na.rm = TRUE ), median_steps_day=median(steps, na.rm = TRUE ))
step_day_stat

summarize(activity, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))
```

## What is the average daily activity pattern?
### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
step_interval = group_by(activity, interval)
step_interval_mean= summarise(step_interval, mean=mean(steps, na.rm= T))
with(step_interval_mean, plot(interval, mean, type="l", main="Average Number of Steps per Interval", xlab="Interval", ylab="Average Number of Steps" ))
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
which.max(step_interval_mean$mean)
step_interval_mean[104,]
```

So interval 835 has the maximum average number of steps.


## Imputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity$steps))
```


### A strategy for filling in all of the missing values.
My strategy is to use the average of that day for the missing valuse, if the average of the day is missing then I use the overall average.

```{r}

total_mean=mean(activity$steps, na.rm =TRUE)
activity_mean=merge(activity, step_day_stat, by=intersect(names(activity), names(step_day_stat)))
activity_mean2=merge(activity, step_day_stat, by=intersect(names(activity), names(step_day_stat)))

for(i in 1:length(activity_mean$steps)){if(is.na(activity_mean[i,5])){activity_mean[i,5]=total_mean}}
for(i in 1:length(activity_mean$steps)){if(is.na(activity_mean[i,2])){activity_mean[i,2]=activity_mean[i,5]}}


```

```{r}
g3=ggplot(activity_mean, aes(day, mean_steps_day))+geom_histogram(stat = "identity", fill="red")+ggtitle("Total Number of Steps per Day (After Filling NAs)")+ xlab("Day")+ylab("Total Number of Steps")
plot(g3)

activity_mean_day=group_by(activity_mean, day)
summarize(activity_mean_day, mean=mean(steps), median=median(steps))
summarize(activity_mean, mean=mean(steps), median=median(steps))
```
Total mean is not changing because my analysis is showing that either the value for all intervals in a day is missing or non of them is missing. This means that with the strategy that I used all missing values are substituted by total mean of the data which is 37.3826. Because of this the total mean after filling the missing values won't change.If you compare figure ... you will see that the messing values are substituted by the average total step in a day which is :37.3826*288= 10,656

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

For this part I used the filled data from previous part
```{r}
activity_mean$wday=weekdays(activity_mean$date)
activity_wday=group_by(activity_mean, wday)
activity_wday_mean=summarize(activity_wday, mean_wday=mean(steps, na.rm =TRUE))
g4=ggplot(activity_wday_mean, aes(wday, mean_wday))+geom_histogram(stat = "identity", fill="blue")+ggtitle("Average Number of Steps for Weekdays")+ xlab("Weekday")+ylab("Average Number of Steps")
plot(g4)
activity_mean$ffday="Weekday"
for(j in 1:length(activity_mean$steps)){if(activity_mean$wday[j]=="Saturday"|activity_mean$wday[j]=="Sunday"){activity_mean$ffday[j]="Weekend"}}


```

We can see that the average number of steps is generally higher during Weekend. 

### Make a panel plot 

```{r}
activity_interval=group_by(activity_mean, ffday, interval)
activity_interval_mean=summarize(activity_interval, mean_interval=mean(steps, na.rm =TRUE))
ggplot(activity_interval_mean, aes(interval, mean_interval))+geom_line(col="blue")+facet_grid(ffday~.)+ggtitle("Average Number of Steps per Interval for Weekdays and Weekend")+ xlab("Interval")+ylab("Average Number of Steps")



```

By comparing number of steps in each interval we can see that it is not the same during weekend and weekdays. However there is no specific trend sometimes it is higher in weekends and sometimes it is lower.
